---
title: "Data and Stylized facts"
author: "Lorenzo Lucchese"
date: "14/02/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(readxl)
library(zoo)
library(xts)
library(forecast)
library(ggplot2)
library(cowplot)
```

## Data and stylized facts

In this thesis we will consider the S&P 500 index from January 2018 to October 2020. The data can be obtained for the SPY ticker (an ETF which tracks the S&P 500 index) from LOBSTER (https://lobsterdata.com/) in the form of high frequency limit order book data. We appropriately preprocess the data to obtain a time series of equidistant high frequency returns by using the code available at https://github.com/lorenzolucchese/LOBSTER-data-processing. These will be used to compute the realised measures in section [add reference]. 

For the moment we focus on daily returns and highlight some important empirical characteristics that can be observed in general financial time series, known as the stylized facts:
\begin{enumerate}[start=1, label={(\bfseries S\arabic*):}]
\item Return series are not identically independently distributed although they show little serial correlation.
\item Series of absolute or squared returns show profound serial correlation.
\item Conditional expected returns are close to zero.
\item Volatility appears to vary over time.
\item Return-series are heavy tailed.
\item Extreme returns appear in clusters, this phenomenon is known as volatility clustering.
\end{enumerate}

```{r}
# Data 
SPY_preprocessed_data <- read_excel(
  "C:/Users/Lorenzo/Desktop/Imperial/M4/Thesis/SPY_1min_preprocessed_data.xlsx",
  col_names = FALSE, col_types = c("date", "numeric"))

corrupted_dates <- c("2019-01-09", "2020-05-26", "2020-06-04", "2020-08-05", "2020-08-10",
                     "2020-08-18", "2020-09-24", "2020-10-30")
SPY.HF <- xts(read.zoo(SPY_preprocessed_data))
# remove corrupted dates
SPY.HF <- SPY.HF[!(time(SPY.HF) %in% index(SPY.HF[corrupted_dates]))]

# daily returns
SPY.dailyreturns <- read_excel(
  "C:/Users/Lorenzo/Desktop/Imperial/M4/Thesis/SPY_daily_log_returns.xlsx",
  col_names = FALSE, col_types = c("date", "numeric"))
SPY.dailyreturns <- xts(read.zoo(SPY.dailyreturns))
# remove corrupted dates
SPY.dailyreturns <- SPY.dailyreturns[!(index(SPY.dailyreturns) %in%
                                       index(SPY.dailyreturns[corrupted_dates]))]

head(SPY.dailyreturns)

p <- ggplot(SPY.dailyreturns, aes(x = Index)) + geom_line(aes(y = x)) +
  ggtitle("SPY daily returns") + ylab("") + xlab("") +
  theme(plot.title = element_text(hjust = 0.5))
p

save_plot("Thesis plots/SPYdailyreturns.jpg", p, dpi = 600, limitsize = FALSE)
```


#### (S1) Return series are not identically independently distributed although they show little serial correlation.

```{r}
# We plot the autocorrelation sequence for the SPY returns (max lag 1 month)
p <- ggAcf(SPY.dailyreturns, lag.max = 22) + ylim(c(-1, 1)) +
  ggtitle("Autocorrelation of returns") + 
  theme(plot.title = element_text(hjust = 0.5))
p
save_plot("Thesis plots/S1acvs.jpg", p, dpi = 600, limitsize = FALSE)
```

#### (S2) Series of absolute or squared returns show profound serial correlation.

```{r}
# We plot the autocorrelation sequence for the squared SPY returns (max lag 1 month)
p <- ggAcf(SPY.dailyreturns^2, lag.max = 22) + ylim(c(-1,1)) +
  ggtitle("Autocorrelation of squared returns") +
  theme(plot.title = element_text(hjust = 0.5))
p
save_plot("Thesis plots/S2acvs.jpg", p, dpi = 600, limitsize = FALSE)
```

#### (S3) Conditional expected returns are close to zero.

```{r}
# [elaborate]
```


#### (S4) Volatility appears to vary over time.

```{r}
# We plot the 1-week rolling volatility.

p <- ggplot(sqrt(rollapply(SPY.dailyreturns^2, 5, mean)), aes(x = Index)) +
  geom_line(aes(y = x)) + xlab("") + ylab("") + 
  ggtitle("1 week rolling volatility") +
  theme(plot.title = element_text(hjust = 0.5))
p

save_plot("Thesis plots/S4rollvol.jpg", p, dpi = 600, limitsize = FALSE)
```


#### (S5) Return-series are heavy tailed.

```{r}
# We plot the empirical return density for the whole period. The density of a 
# normal distribution with fitted mean and standard deviation is plotted for 
# comparison.

p <- ggplot(SPY.dailyreturns) + 
  geom_density(aes(x = x, col = "Empirical density"), adjust = 2) +
  stat_function(fun = dnorm, n = 1001, args = 
                  list(mean = mean(SPY.dailyreturns), sd = sd(SPY.dailyreturns)),
                aes(colour = "Normal density")) +
  coord_cartesian(xlim = c(-0.04, 0.04)) + 
  xlab("Daily returns") + ylab("") +
  scale_color_manual(values = c("black", "red")) + 
  guides(colour = "legend") +
  ggtitle("Empirical return density") +
  theme(legend.position = c(0.85,0.8), legend.title = element_blank(), 
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        plot.title = element_text(hjust = 0.5))
p 
p_zoom <- p + coord_cartesian(xlim = c(-0.06, -0.03), ylim = c(0, 2.5)) + 
  ggtitle("Empirical return density (lower tail)")
p_zoom

p_QQ <- ggplot(SPY.dailyreturns, aes(sample = x)) + stat_qq() + stat_qq_line() +
  ggtitle("Normal quantile-quantile plot of returns") +
  theme(plot.title = element_text(hjust = 0.5))
p_QQ

p <- plot_grid(p, p_zoom, p_QQ, nrow = 3)

p

save_plot("Thesis plots/S5fattails.jpg", p, nrow = 3, dpi = 600, limitsize = FALSE)
```

#### (S6) Extreme returns appear in clusters, this phenomenon is known as volatility clustering.
```{r}
# We plot only the returns with absolut value bigger than the 95-th quantile
highquantile <- quantile(abs(SPY.dailyreturns), 0.95)
SPY.extremereturns <- SPY.dailyreturns[abs(SPY.dailyreturns) > highquantile]

p <- ggplot(SPY.dailyreturns, aes(x = Index)) +
  geom_line(aes(y = x)) +
  annotate("point", x = as.POSIXct(index(SPY.extremereturns)), 
           y = as.numeric(SPY.extremereturns), col = "red") +
  xlab("") + ylab("") + ggtitle("SPY extreme returns") +
  theme(plot.title = element_text(hjust = 0.5))
p

save_plot("Thesis plots/S6clustering.jpg", p, dpi = 600, limitsize = FALSE)

```


## Measuring tail risk: VaR and ES

The following plot helps to visualise the concepts of VaR and ES.

```{r}
alpha <- 0.05
DF <- 5
density_df <- data.frame(x = seq(-5, 5, 0.001), 
                         y = dt(seq(-5, 5, 0.001), df = DF, ncp = 0))
VaR <- qt(1-alpha, df = DF, ncp = 0)
ES <- integrate(function(gamma) qt(gamma, df = DF, ncp = 0), 1-alpha, 1)$value / alpha
p <- ggplot(density_df, aes(x)) + 
  geom_line(aes(x= x, y= y)) +
  geom_vline(linetype = "dotdash", xintercept = VaR, col ="darkred") +
  geom_vline(linetype = "dotdash", xintercept = ES, col = "darkred") +
  geom_area(data = density_df[density_df$x > VaR,], aes(y = y), alpha = 0.2) +
  coord_cartesian(c(-2, 4)) + xlab("Loss distribution, L = -Y") + ylab("") +
  ggtitle("VaR and ES visualised") +
  theme(plot.title = element_text(hjust = 0.5)) +
  annotate("text", x = VaR, y = -0.01, 
             label = paste0(round(100*(1-alpha)),"% VaR")) +
  annotate("text", x = ES, y = -0.01, label = paste0(round(100*(1-alpha)),"% ES")) 

p

save_plot("Thesis plots/VaRES.jpg", p, dpi = 600, limitsize = FALSE)
```

### Shortcomings of VaR and ES

Different tails given rise to the same (VaR, ES) couple.

```{r}
# We use alpha and (VaR, ES) from the previous distribution

# slowly decaying tail: (1+x)^-alpha
gamma <- 2 + 1/(ES-VaR)

# exponentially decaying tail: exp(-beta*x)
beta <- 1/(ES-VaR)

# finite support tail (1+xi*x)^(-1-1/xi), for x <= -1/xi and xi < 0
xi <- 1 - 1/(ES-VaR)

x <- seq(VaR - 1, -1/xi + 0.5, 0.001)
tails_df <- data.frame(x, slow.tail <- alpha/(gamma-2)*(1+x)^(-gamma),
                       exp.tail <- alpha*beta*exp(-beta*x),
                       finite.tail <- c(alpha*(1+xi*x[x<(-1/xi)])^(-1-1/xi), 
                                        rep(0, sum(x>=(-1/xi)))))

p <- ggplot(tails_df, aes(x = x)) +
  geom_line(aes(y = slow.tail, linetype = "Slowly varying")) +
  geom_line(aes(y = exp.tail, linetype = "Exponential")) +
  geom_line(aes(y = finite.tail, linetype = "Finite support")) + 
  geom_vline(xintercept = VaR, linetype = "dotdash", col = "darkred") +
  geom_vline(xintercept = ES, linetype = "dotdash", col = "darkred") +
  ggtitle("Different tails, same VaR and ES") +
  theme(legend.position = c(0.8, 0.8), plot.title = element_text(hjust = 0.5)) + 
  coord_cartesian(xlim = c(VaR - 0.5, -1/xi+0.5), ylim = c(0, 0.0075)) +
  scale_linetype_manual("Tail behaviour", 
                        breaks = c("Slowly varying", "Exponential", 
                                   "Finite support"),
                        values = c("dashed", "longdash", "solid")) + 
  ylab("") + xlab("Loss distribution, L = -Y") +
  annotate("text", x = VaR, y = 0, 
             label = paste0(round(100*(1-alpha)),"% VaR")) +
  annotate("text", x = ES, y = 0, label = paste0(round(100*(1-alpha)),"% ES")) 

p

p_zoom <- p + coord_cartesian(xlim = c(4, -1/xi+0.5), ylim = c(0, 0.001)) +
  ggtitle("Different tails, same VaR and ES (zoomed)") + 
  theme(legend.position = "none")

p <- plot_grid(p, p_zoom, nrow = 2)

p

save_plot("Thesis plots/VaRESshortcomings.jpg", p, nrow = 2, dpi = 600, limitsize = FALSE)
```

