---
title: "Data and Stylized facts"
author: "Lorenzo Lucchese"
date: "14/02/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(readxl)
library(zoo)
library(xts)
library(forecast)
library(rugarch)
library(ggplot2)
```

## RV stylized facts

By fixing a 1-day time period ($T = 1$) and a high-frequency time interval $\Delta$ (usually 1 or 5 minutes) we define the associated realised variance $(RV_t)_{t\in\mathbb{N}}$ process for the return process $(Y_t)_{t\in\mathbb{N}}$ s.t. $RV_t$ is the realised variance of the price process over the time period $[t-1, t]$ with $\Delta$ sampling, i.e. the sum of the squared $\Delta$-frequency returns between time $t-1$ and $t$. We can hence look back at the S&P 500 return data introduced in section [add reference] and point out the following additional well-known empirical stylized facts of return series:
\begin{enumerate}[start=7, label={(\bfseries S\arabic*):}]
\item realised variance displays long memory: its autocorrelation series has hyperbolic rate of decay, i.e. $\rho_t(RV) \sim t^{\gamma}$ for $\gamma\in (0,1)$.
\item The asset return process displays multi-scaling behaviour, which in terms of "generalised volatilities" of the (continuous-time) asset return process implies $\mathbb{E}[X_t^q]\sim t^{\tau(q)+1}$ where $\tau(\cdot)$ is a concave scaling function. 
\item The return series display volatility cascades, i.e. volatility over longer time intervals has stronger influence on that at shorter time intervals than conversely.
\end{enumerate}

\textbf{Remark} We define a (mono)-scaling process with index $H$ by a (continuous-time) stochastic process $X$ s.t. $X_0 = 0$ and for any $c>0$ and sequence of times $t_1,t_2,\ldots, t_k$
\[(X_{t_1}, X_{t_2}, \ldots, X_{t_k}) \overset{d}{=} c^{-H}(X_{ct_1}, X_{ct_2}, \ldots, X_{ct_k}).\]
A mono-scaling process has the property that $\mathbb{E}[X_t^q] = c(q) t^{Hq}$ where $c(\cdots)$ is a deterministic function of $q$. We can generalise this idea and define a multi-scaling process $X$ as a continuous-time stochastic process with stationary increments s.t. $\mathbb{E}[X_t^q] =c(q) t^{\tau(q)+1}$ for all $q\in\mathcal{Q}$ with $[0,1]\subseteq\mathcal{Q}$. We refer to [add reference] (Mandelbrot) for an in-depth introduction to multi-scaling processes, and note that in this paper the authors actually propose a model for asset returns purely motivated by this property.

```{r}
# Data 
SPY_preprocessed_data <- read_excel(
  "C:/Users/Lorenzo/Desktop/Imperial/M4/Thesis/SPY_1min_preprocessed_data.xlsx",
  col_names = FALSE, col_types = c("date", "numeric"))

corrupted_dates <- c("2019-01-09", "2020-05-26", "2020-06-04", "2020-08-05", "2020-08-10",
                     "2020-08-18", "2020-09-24", "2020-10-30")
SPY.HF <- xts(read.zoo(SPY_preprocessed_data))
# remove corrupted dates
SPY.HF <- SPY.HF[!(time(SPY.HF) %in% index(SPY.HF[corrupted_dates]))]

# daily returns
SPY.dailyreturns <- read_excel(
  "C:/Users/Lorenzo/Desktop/Imperial/M4/Thesis/SPY_daily_log_returns.xlsx",
  col_names = FALSE, col_types = c("date", "numeric"))
SPY.dailyreturns <- xts(read.zoo(SPY.dailyreturns))
# remove corrupted dates
SPY.dailyreturns <- SPY.dailyreturns[!(index(SPY.dailyreturns) %in%
                                       index(SPY.dailyreturns[corrupted_dates]))]
# compute realised variance
SPY.RV <- apply.daily(SPY.HF, function(x) sum(x^2))
head(SPY.RV)

p <- ggplot(SPY.RV, aes(x = Index)) + geom_line(aes(y = x)) +
  ggtitle("SPY realised variance") + ylab("") + xlab("") +
  theme(plot.title = element_text(hjust = 0.5))
p

save_plot("Thesis plots/SPYrealisedvariance.jpg", p, dpi = 600, limitsize = FALSE)
```

#### (S7) Long memory in volatility
```{r}
# We plot the autocorrelation sequence for the SPY returns (max lag 1 month)
p <- ggAcf(sqrt(SPY.RV), lag.max = 30) +
  ggtitle("Autocorrelation of realised volatility") + 
  theme(plot.title = element_text(hjust = 0.5))
p

save_plot("Thesis plots/S7acvs.jpg", p, dpi = 600, limitsize = FALSE)

```

#### (S8) The asset return process displays multi-scaling behaviour.


#### (S9) Asymmetric propagation of volatility (volatility cascade effect).


### Comparing the acvs of RV, GARCH and HAR

```{r}
RV.acf <- acf(sqrt(SPY.RV), lag.max = 40, plot = FALSE)$acf

fit.spec <- ugarchspec(variance.model = list(model = "sGARCH", garchOrder = c(1, 1)),
                       mean.model = list(armaOrder = c(0, 0), include.mean = FALSE),
                       distribution.model = "norm")
GARCH.fit <- ugarchfit(SPY.dailyreturns, spec = fit.spec)
GARCH.acf <- acf(sigma(GARCH.fit), lag.max = 40, plot = FALSE)$acf

HAR.fit <- HARmodel(SPY.RV)
HAR.acf <- acf(sqrt(HAR.fit$fitted.values), lag.max = 40, plot = FALSE)$acf

acf.df <- data.frame(lag = 1:41, RV.acf, GARCH.acf, HAR.acf)

p <- ggplot(acf.df, aes(x = lag)) + 
  geom_line(aes(y = predict(loess(RV.acf ~ lag, data = acf.df)), lty = "Observed")) +
  geom_line(aes(y = predict(loess(GARCH.acf ~ lag, data = acf.df)), lty = "GARCH(1, 1)")) +
  geom_line(aes(y = predict(loess(HAR.acf ~ lag, data = acf.df)), lty = "HAR")) +
  labs(lty = "Model") + ylab("ACF") +
  ggtitle("ACFs of observed realised volatility and model fitted values") +
  theme(plot.title = element_text(hjust = 0.5), legend.position = c(0.8,0.8))
p

save_plot("Thesis plots/GARCHvsHAR.jpg", p, dpi = 600, limitsize = FALSE)
```

