---
title: "Tail Risk Management - CPOT simulation"
author: "Lorenzo Lucchese"
date: "07/03/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(rugarch)
library(zoo)
library(kableExtra)
library(xts)
library(cowplot)
library(grid)
library(gridExtra)
library(highfrequency)
```

## C-POT model
We model the tails of the time series $\{X_t, t\in\mathbb{Z}\}$ conditional on the filtration $(\mathcal{F}_t)_t$ by:
\begin{align*}
	\mathbb{P}(X_t>x|\mathcal{F}_{t-1}) &= \mathbb{P}(X_t>x|X_t\geq u, \mathcal{F}_{t-1})\mathbb{P}(X_t\geq u|\mathcal{F}_{t-1}) = \\
	& = (1-F^u_{t|t-1}(x-u))(1-F_{t|t-1}(u)) = \\
	& \approx \phi_{t|t-1} \Big(1 + \xi_{t|t-1} \frac{x-u}{\beta_{t|t-1}}\Big)^{-1/\xi_{t|t-1}}.
\end{align*}

We specify separate dynamic models for $\phi_{t|t-1}$ and GPD parameters $\xi_{t|t-1},\beta_{t|t-1}$:
\begin{itemize}
	\item We consider a Logit Binomial GLM for the exceedances, $I_t=1_{\{X_t\geq u\}}$, with covariates $\boldsymbol{RM}_{t-1}$:
	\begin{align*}
		\mathbb{P}(I_t=i|\mathcal{F}_{t-1}) &= \phi_{t|t-1}^i(1-\phi_{t|t-1})^{1-i} \qquad i\in\{0,1\}\\
		\phi_{t|t-1} &= \textrm{logit}^{-1}(\boldsymbol{\psi}^T\boldsymbol{RM}_{t-1})
	\end{align*}
	i.e. $I_t|\mathcal{F}_{t-1}\sim Bernoulli(\phi_{t|t-1})$ where $\boldsymbol{\psi} = [\psi_0,\psi_1,\ldots, \psi_p]$ and $\boldsymbol{RM}_{t-1} = [1,RM_{t-1}^1,\ldots, RM_{t-1}^p]$.
	
	\item We model $F^u_{t|t-1}$ as a GPD distribution with time varying shape and scale parameters 
	\begin{align*}
		\xi_{t|t-1} &= \exp(\boldsymbol{RM}_{t-1}^T\boldsymbol{\nu}) \\
		\beta_{t|t-1} &= \exp(\boldsymbol{RM}_{t-1}^T\boldsymbol{\kappa})
	\end{align*}
	where $\boldsymbol{\nu} = [\nu_0,\nu_1,\ldots, \nu_p]$, $\boldsymbol{\kappa} = [\kappa_0,\kappa_1,\ldots, \kappa_p]$ and $\boldsymbol{RM}_{t-1} = [1,RM_{t-1}^1,\ldots, RM_{t-1}^p]$. Here we are hence modelling $X_t-u|\mathcal{F}_{t-1},I_t=1$ as a GPD distribution with parameters $\xi_{t|t-1}$ and $\beta_{t|t-1}$.
\end{itemize}

The MLEs for the parameters are given by:
\begin{align*}
	\boldsymbol{\hat\psi} &=  \underset{\boldsymbol{\psi}\in \Psi}{\text{argmax}} \frac{1}{n} \sum_{t=1}^n\bigg\{ i_t\log\big(\textrm{logit}^{-1}(\boldsymbol{\psi^T\boldsymbol{RM}_{t-1}})\big) + (1-i_t)\log\big(1-\textrm{logit}^{-1}(\boldsymbol{\psi}^T\boldsymbol{RM}_{t-1})\big)\bigg\}\\
	(\boldsymbol{\hat\nu},\boldsymbol{\hat\kappa}) &= \underset{(\boldsymbol{\nu},\boldsymbol{\kappa})\in N\times K}{\text{argmax}} \frac{1}{n} \sum_{t=1}^n \bigg\{i_t \log\Big(g_{\exp\{\boldsymbol{\nu}^T\boldsymbol{RM}_{t-1}\}), \exp\{\boldsymbol{\kappa}^T\boldsymbol{RM}_{t-1}\}}(x_t)\Big)\bigg\}
\end{align*}
We note that the first estimator is the MLE of a Binomial GLM with logit link function.


```{r}
GPD_loglikelihood_CPOT <- function(x, RM, nu, kappa){
  #================================================================================
  # A function which computes the log-likelihood of a sample from a GPD distribution 
  # with shapes and scales depending on the covariates RM through:
  # xi[i] = exp( RM[i,]^T nu ) and beta[i] = exp( RM[i,]^T kappa  )
  #
  # INPUTS
  # x: the GPD sample. (n x 1)
  # RM: the covariates matrix (xi[i], beta[i] will be "regressed" on RM[i,]). (n x p)
  # nu: the parameters determining the shape of the GPD distribution. (p x 1)
  # kappa: the parameters determining the scale of the GPD distribution. (p x 1)
  #
  # OUTPUTS
  # loglik: the log likelihood evaluated at x. (1 x 1)
  #================================================================================
  xi <- exp( RM %*% nu )
  beta <- exp( RM %*% kappa )
  # by specification of xi and beta we have xi, beta > 0
  loglik <- sum((-1/xi-1)*log(1+xi*x/beta) - log(beta))
  return(loglik)
}

GPD_loglikelihood_CPOT_gr <- function(x, RM, nu, kappa){
  #================================================================================
  # A function which computes the gradient of the log-likelihood of a sample from a 
  # GPD distribution with shapes and scales depending on the covariates RM through:
  # xi = exp( RM[t,]^T nu ) and beta = exp( RM[t,]^T kappa  )
  #
  # INPUTS
  # x: the GPD sample. (n x 1)
  # RM: the covariates matrix (x[t] will be "regressed" on RM[t,]). (n x p)
  # nu: the parameters determining the shape of the GPD distribution. (p x 1)
  # kappa: the parameters determining the scale of the GPD distribution. (p x 1)
  #
  # OUTPUTS
  # loglik_gr: the log likelihood evaluated at x. (1 x 1)
  #================================================================================
  xi <- exp( RM %*% nu )
  beta <- exp( RM %*% kappa )
  # by specification of xi and beta we have xi, beta > 0, the derivatives are given by:
  grad_loglik_nu <- t(1/xi*log(1 + xi/beta * x) - (xi + 1) * x/(beta + xi*x))%*%RM
  grad_loglik_kappa <- t( (xi + 1) * x / (beta + xi * x) - 1 )%*%RM
  loglik_gr <- c(grad_loglik_nu, grad_loglik_kappa)
  return(loglik_gr)
}


CPOT_fit <- function(x, u, RM, IID.covariates = TRUE){
  #================================================================================
  # A function which fits the C-POT model to the sample x above the specified 
  # threshold u with covariates RM. The MLE estimators for psi, nu and kappa are 
  # computed. If IID.covariates = TRUE (default) the asymptotic standard errors 
  # for the parameters are computed, along with the asymptotic variance covariance
  # matrix for the nu, kappa MLEs
  #
  # INPUTS
  # x: the data sample. (n x 1)
  # RM: the covariates matrix (x[i] will be "regressed" on RM[i,]). (n x p)
  # u: the threshold u above which to fit the GPD distribution. (1 x 1)
  # IID.covariates: logical indicating whether the covariates can be treated as IID.
  #
  # OUTPUTS
  # CPOT_fit: a list with named components:
  #           pars: a dataframe with columns psi, nu, kappa corresponding to the 
  #                 compute MLEs. (p x 3)
  #           pars_sd: a dataframe with columns psi_se, nu_se, kappa_se corresponding 
  #                    to the asymptotic standard errors of the MLEs when covariates
  #                    are IID. (p x 3)
  #           nukappa_varcov: asymptotic variance covariance matrix for the nu and 
  #                           kappa MLEs when the covariates are IID. (2p x 2p)
  #================================================================================
  n <- length(x)
  p <- ncol(RM)
  
  # if there are not enough exceedances to fit a GPD distribution return error
  if(sum(x>u) <= 2) stop("Too few exceedances: lower threshold u")
  
  # compute MLE for psi using glm function
  i <- as.numeric(x > u)
  logitGLM_data <- data.frame(i, RM)
  logitGLM <- summary(glm(i ~ 0 + ., family=binomial(link="logit"), data = logitGLM_data))
  psi <- logitGLM$coefficients[,1]
  
  # compute MLE and standard error for GPD distribution on exceedances
  nu_kappa <- optim(rep(0, (2*p)), 
                   fn = function(pars) 
                     - GPD_loglikelihood_CPOT(x[x>u] - u, RM[(x>u),], pars[1:p], 
                                              pars[(p+1):(2*p)]),
                   gr = function(pars) 
                     - GPD_loglikelihood_CPOT_gr(x[x>u] - u, RM[(x>u),], pars[1:p],
                                                 pars[(p+1):(2*p)]))
  nu <- nu_kappa$par[1:p]
  kappa <- nu_kappa$par[(p+1):(2*p)]
  
  # if the regressors RM are IID compute the analytic asymptotic standard errors
  if(IID.covariates){
    # for psi
    psi_se <- logitGLM$coefficients[,2]
    
    # for nu, kappa
    xi <- exp( RM %*% nu)
    J <- 0
    for(t in which(x>u)){
      M <- matrix(rep(0, (2*p)^2), nrow = (2*p))
      M[1:p, 1:p] <- 2 * xi[t]^2 * RM[t,]%*%t(RM[t,])
      M[1:p, (p+1): (2*p)] <- xi[t] * RM[t,]%*%t(RM[t,])
      M[(p+1):(2*p), 1:p] <- xi[t] * RM[t,]%*%t(RM[t,])
      M[(p+1):(2*p), (p+1):(2*p)] <- (xi[t] + 1) * RM[t,]%*%t(RM[t,])
      M <- 1/((xi[t]+1) * (2*x[t]+1) * sum(i)) * M
      J <- J +  M
    }

    # asymptotic sd of xi.hat, kappa.hat
    varcov <- solve(sum(i) * J)
    nu_se <- sqrt(diag(varcov))[1:p]
    kappa_se <- sqrt(diag(varcov))[(p+1):(2*p)]
  }
  # if the covariates cannot be taken as IID
  else{
    # do not compute standard errors
    psi_se <- NA
    nu_se <- NA
    kappa_se <- NA
    varcov <- NA
  }
  
  
  # return results
  return(list(pars = data.frame(psi, nu, kappa),
              pars_se = data.frame(psi_se, nu_se, kappa_se),
              nukappa_varcov = varcov))
  
}

```

```{r}
CPOT_bootstrap <- function(x, u, RM, pars){
  #================================================================================
  # A function which creates a bootstrap resample of the data for fitted values of 
  # the parameters psi, nu, kappa.
  # 
  # INPUTS
  # x: the original data sample. (n x 1)
  # u: the C-POT threshold. (1 x 1)
  # RM: the C-POT covariates. (n x p)
  # pars: the fitted parameters, a dataframe with columns psi, nu, kappa. (p x 3)
  #
  # OUTPUTS
  # x_: a resample of the data via bootstrap. (n x 1)
  #================================================================================
  psi <- pars$psi
  nu <- pars$nu
  kappa <- pars$kappa
  
  # fitted parameters
  phi <- exp( RM%*%psi )/(1 + exp( RM%*%psi ))
  xi <- exp( RM%*%nu )
  beta <- exp( RM%*%kappa )
  
  # compute exceedence residuals
  r <- 1/xi[x>u] * log(1 + xi[x>u]/beta[x>u] * (x[x>u]-u) )
  
  n <- nrow(RM)
  x_ <- rep(0, n)
  
  # simulate exceedances
  i <- (runif(n) < phi )
  # if exceedance has occured simulate exceedance residual and post-blacken
  x_[i] <- u + beta[i]/xi[i] * ( exp(xi[i]*sample(r, sum(i), replace = TRUE)) - 1)
  # if exceedance has not occured simulate from the empirical distribution 
  # function of values below u
  x_[!i] <- sample(x[x<=u], sum(!i), replace = TRUE) 
  
  return(x_)
}

CPOT_bootstrapCIs <- function(x, u, RM, conf.level = 0.95, boot = 500){
  #================================================================================
  # A function which computes the bootstrap confidence intervals for the parameters 
  # of the C-POT model with data x and covariates RM.
  # 
  # INPUTS
  # x: the original data sample. (n x 1)
  # u: the C-POT threshold. (1 x 1)
  # RM: the C-POT covariates. (n x p)
  # conf.level: the confidence level of the CIs. (1 x 1)
  # boot: number of bootrap samples to use. (1 x 1)
  #
  # OUTPUTS
  # pars_CIs: a list with named components:
  #           psi, nu, kappa: dataframes containing the CIs (2 x p).
  #================================================================================
  # fit the model
  pars <- CPOT_fit(x, u, RM, IID.covariates = FALSE)$pars
  p <- ncol(RM)
  
  # initialise matrices to contain bootstrap samples (each row corresponds to a new sample)
  psi_boot <- matrix(rep(0, p*boot), nrow = boot)
  kappa_boot <- matrix(rep(0, p*boot), nrow = boot)
  nu_boot <- matrix(rep(0, p*boot), nrow = boot)
  
  for(b in 1:boot){
    # form a new bootstrap sample
    x_ <- CPOT_bootstrap(x, u, RM, pars)
    
    # fit the model to the new sample
    pars_boot <- CPOT_fit(x_, u, RM, IID.covariates = FALSE)$pars
    
    # retain the fitted parameters
    psi_boot[b,] <- pars_boot$psi
    nu_boot[b,] <- pars_boot$nu
    kappa_boot[b,] <- pars_boot$kappa
  }
  
  # probability
  alpha <- 1 - conf.level
  
  # compute the confidence intervals using the empirical distribution for the parameters
  # obtain through bootstrapping
  psi_CI <- apply(psi_boot, 2, function(x) quantile(x, probs = c(alpha/2, 1-alpha/2)))
  nu_CI <- apply(nu_boot, 2, function(x) quantile(x, probs = c(alpha/2, 1-alpha/2)))
  kappa_CI <- apply(kappa_boot, 2, function(x) quantile(x, probs = c(alpha/2, 1-alpha/2)))
  
  # return the bootstrap confidence intervals
  return(list(psi_CI = psi_CI, nu_CI = nu_CI, kappa_CI = kappa_CI))
}
```

We carry out a data simulation study to check the CPOT functions are working properly. We first simulate a sample $\boldsymbol{x} = (x_t)_{t=1}^n$ from the model-assumed data generating process with IID covariates $\boldsymbol{RM}_t$ s.t. $RM^{(1)}\sim N(0,1)$, $RM^{(2)}\sim N(0,1)$ independently. 
We fix  $\boldsymbol{\psi}, \boldsymbol{\nu}, \boldsymbol{\kappa}$ and simulate exceedance indicators $i_t \sim Bernoulli(\text{logit}^{-1}(\boldsymbol{\psi}^T\boldsymbol{RM}_{t-1}))$ and exceedances $x_t - u \sim GPD(\exp(\boldsymbol{\nu}^T\boldsymbol{RM}_{t-1}), \exp(\boldsymbol{\kappa}^T\boldsymbol{RM}_{t-1}))$. For non-exceedances, i.e. $i_t = 0$, we pick an arbitrarily distribution with support in $(-\infty, u)$.

We fit the model using _CPOT_fit()_ and compare the MLEs with the true parameters.
We also compute the confidence intervals for the parameter estimates using both the closed form asymptotic results (which are applicable since the covaraites are IID) and the bootstrapping technique. We check the two methods yield similar values. All results are displayed in the tables below.

\textbf{Remark} This is also a good starting point for a detailed analysis of the parameter MSEs and the CI widths produced by the C-POT specification.

```{r}
# Data simulation study
n <- 5000
p <- 2
u <- 0.5
psi <- c(-5, -1)
nu <- c(0.1, 0.2)
kappa <- c(0.05, 0.5)
RM <- cbind(rnorm(n), rnorm(n, 2))
pars <- data.frame(psi, nu, kappa)

CPOT_sim <- function(RM, u, pars){
  #================================================================================
  # A function which simulates a sample from the CPOT model specification with 
  # covariates RM.
  #
  # INPUTS
  # RM: model covariates. (n x p)
  # u: threshold. (1 x 1)
  # pars: A dataframe with columns psi, nu, kappa. (p x 3)
  #
  # OUTPUTS
  # x: sample from C-POT model. (n x 1)
  #================================================================================
  psi <- pars$psi
  nu <- pars$nu
  kappa <- pars$kappa
  
  # time-varying parameters
  phi <- exp( RM%*%psi )/(1 + exp( RM%*%psi ))
  xi <- exp( RM%*%nu )
  beta <- exp( RM%*%kappa )
  
  # initialise sample
  n <- nrow(RM)
  x <- rep(0, n)
  
  # simulate exceedances
  i <- (runif(n) < phi)
  #if exceedance has occured simulate from GPD (upper tail)
  x[i] <- u + beta[i]/xi[i] * ( exp(xi[i]*rexp(sum(i))) - 1)
  # if exceedance has not occured simulate from any non parametric distribution
  # with support on (-infty, u) - alternatively set to 0
  x[!i] <- runif(sum(!i), u - 2, u) 
  
  return(x)
}

x <- CPOT_sim(RM, u, pars)

# check model fitting 
C <- CPOT_fit(x, u, RM)
kable(pars, format = "simple",
      caption = "True parameters",
      escape = F, digits = 4)
rownames(C$pars) <- NULL
kable(C$pars, format = "simple",
      caption = "Fitted parameters",
      escape = F, digits = 4)

```

```{r}
# Check bootstrapping and asymptotic CIs yield similar values (asymptotically the same)

# Asymptotic confidence intervals
psi_CI <- data.frame(rbind(C$pars$psi - 1.96*C$pars_se$psi, 
                           C$pars$psi + 1.96*C$pars_se$psi))
nu_CI <- data.frame(rbind(C$pars$nu - 1.96*C$pars_se$nu, 
                          C$pars$nu + 1.96*C$pars_se$nu))
kappa_CI <- data.frame(rbind(C$pars$kappa - 1.96*C$pars_se$kappa,
                             C$pars$kappa + 1.96*C$pars_se$kappa))
CI_df <- cbind(psi_CI, nu_CI, kappa_CI)
kable(CI_df, format = "simple",
      caption = "Asymptotic confidence intervals",
      col.names = c(paste0("psi", c(1:p)), paste0("nu", c(1:p)), 
                    paste0("kappa", c(1:p))),
      escape = F, digits = 4)

# Bootstrap confidence intervals
CI_boot <- data.frame(CPOT_bootstrapCIs(x, u, RM))

kable(CI_boot, format = "simple",
      col.names = c(paste0("psi", c(1:p)), paste0("nu", c(1:p)), 
                    paste0("kappa", c(1:p))),
      caption = "Bootstrap confidence intervals",
      escape = F, digits = 4)


```